### 2019-05-16 2-Chapters/2-Chapter/chapter2.tex
Done for today

\begin{figure}[h!]  % [htbp]
    \centering
    \includegraphics[width=0.95\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem__step100.png}
    \caption{Screenshot of the demonstration, after the end of the game, after $t=T=100$ steps.  See \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}}
    \label{fig:2:example_of_a_5_arm_bandit_problem__step100}
\end{figure}

@@

In the example above, we consider Bernoulli distributions, but other distributions have been studied in the literature.
From now on until the last chapter of this thesis, we only consider stochastic rewards. The piece-wise stationary model is studied in Chapter~\ref{chapter:6}.
%
% Everything is implemented and documented at
% https://smpybandits.github.io/docs/Arms.html
We also focus only on real-valued rewards, meaning that $r_k(t)\in\R$ for all $k$ and $t$.
An important hypothesis is whether rewards are bounded or not,
and whether the player knows if they are bounded or not before starting the bandit game.
Moreover, if rewards are known to be bounded, let say in an interval $[a,b]$, another important hypothesis is whether the player knows the values of $a$ and $b$.
%
Intuitively, the bandit game is easier if the player knows the support of the distributions, and we restrict to this case in all the thesis.
Most of the algorithms proposed in the literature follow this hypothesis as well.
%
The mostly used
infinitely supported distributions are Gaussian, exponential and Poisson,
while in the literature, the Bernoulli distribution is the most common case of finitely supported distributions.
Continuous distributions with finite support also included truncated versions of infinitely supported distributions, in particular truncated Gaussian are used for numerical experiments in lots of research articles.

@@

Given a measure $\lambda$, that is usually the natural Lebesgue measure on $\R$, an exponential family of probability distributions is defined as the distributions whose density (relative to $\lambda$) have the following form:
\begin{equation}\label{eq:2:exponentialFamily}
    \Pr(x | \lambda) = h(x) \exp \let( \eta^T T(x) - A(\eta) \right),
\end{equation}
for a parameter vector $\eta$ (the canonical parameter), and given functions $T$ (the sufficient statistic) and $h$.
$A(\eta)$ is the cumulant function, and it is entirely determined by the $\eta$, $h$, $T$,
$A(\eta) = \log \left( \int h(x) \exp(\eta^T T(x)) \lambda(\mathrm{d} x) \right)$.
%
The natural parameter space is the set of values of $\eta$ such that this integral $A(\eta)$ is finite,
and usually the literature focusses on regular and minimal exponential families (when the natural parameter space is an non-empty open set in $\R$).

An important result on exponential families is that a distribution in such family is entirely characterized by its parameter $\eta$.
That is why such one-dimensional distributions are entirely characterized by their mean, as $\mu = \E_{\eta}[T(X)]$.

A second important result is a simplified form for the Kullback-Leibler divergence \cite{KullbackLeibler51}, for two distributions lying in the same exponential family.
The KL divergence between two distributions $d_1=\Pr(x|\eta_1)$ and $d_2=\Pr(x|\eta_2)$ is defined as
\begin{equation}\label{eq:2:Kullback-LeiblerDivergenceExpFamily1}
    \KL\left(\Pr(x|\eta_1), \Pr(x|\eta_2)\right) = \int \Pr(x|\eta_1) \log(\frac{\Pr(x|\eta_1)}{\Pr(x|\eta_1)}) \lambda(\mathrm{d}x) = \E_{d_1}[ \log(\frac{\Pr(x|\eta_1)}{\Pr(x|\eta_1)}) ].
\end{equation}
%
So this KL divergence can be written $\KL( \eta_1, \eta_2 ) = \KL(\Pr(x|\eta_1), \Pr(x|\eta_2))$, and it can be simplified to an expression using only the two parameters $\eta_1$ and $\eta_2$,
and $\mu_1 = \E_{\eta_1}[T(X)]$ the mean of distribution $d_1$:
%
\begin{equation}\label{eq:2:Kullback-LeiblerDivergenceExpFamily2}
    \KL\left( \eta_1, \eta_2 \right) = \E_{d_1}[ (\eta_1 - \eta_2)^T T(X) ] - A(\eta_1) + A(\eta_2) = (\eta_1 - \eta_2)^T \mu_1 - A(\eta_1) + A(\eta_2),
\end{equation}

@@

\paragraph{Other kind of hypothesis: sub-Gaussian and sub-Bernoulli distributions.}
%
A lot of research works considers rewards distributions that are not Gaussian but sub-Gaussian, of a known variance, for instance $1/4$.
For instance, bounded distributions on $[0,1]$ are $1/4$ sub-Gaussian, and this fact is used for instance in \cite{Maillard2018GLR}.
It means that their moment generating function is dominated by that of a Gaussian distribution with the same mean and a variance $1/4$).
In Chapter~\ref{chapter:6}, we instead consider sub-Bernoulli distributions, formally introduced in Definition~\ref{def:6:subBernoulliDistributions}.
%
Such hypothesis is also proposed for other families of distributions, for instance \cite{KimTewari2019} analyses their Follow-the-Perturbed-Leader algorithm for perturbations following any sub-Weibull distribution (sub-Weibull distributions generalize both sub-Gaussian and sub-Exponential distributions).

@@

\paragraph{Application for Opportunistic Spectrum Access.}
% Detail the previous work from our team SCEE on bandits + OSA : Wassim, Navik
%
Previous works of our SCEE team showed that MAB can be used to model the problem of spectrum access for a secondary user accessing a licensed spectrum.
In this model, arms represent a finite set of orthogonal channels, \ie, different frequency bands in a licensed spectrum.
In the model with sensing, the rewards $r_k(t)$ represents the feedback obtained by the CR-equipped device after sensing the channel $k$ at time $t$.
A reward of $r_k(t) = 1$ indicates that no Primary User was sensed, while a reward of $r_k(t)=0$ indicates that the channel $k$ is busy at time $t$ and no uplink message can be sent.
%
This model was first introduced by Wassim Jouini during his PhD thesis,
in \cite{Jouini09} and later studied in both \cite{Jouini10,Jouini12}.
Proof-of-concepts using real-world radio hardware was first propsed in \cite{MoyWSR2014,RobertSDR2014}.
In a second PhD thesis supervised by Christophe Moy, Navikkumar Modi studied the impact
of using MAB algorithms to optimize channel selection
on the battery life of a wireless device.
On the one hand, running a MAB algorithm such as UCB-like algorithms was proved to be useful and can bring significant improvement in terms of successful transmission rates, directly increasing the battery life of the device.
On the other hand, classical MAB algorithms tend to switch arms a lot of times, especially in the beginning of the learning process, and this induces a lot of dynamical hardware reconfiguration for the wireless device, as selecting a different channel requires a change in the radio hardware used by the device.
Each hardware reconfiguration costs energy for the device, and quickly switching algorithms will lead to a reduction of the battery life.
The tradeoff between the two aspects is studied empirically in
\cite{modiDemo2016}.
%
Another interesting works is \cite{Modi17QoS}.

\TODOL{FIXME Finir ce blabla inutile ? Ou alors ce sera déjà fait en intro/chapitre1 ??}

% ----------------------------------------------------------------------------
\section{Lower and upper bounds on regret}
\label{sec:2:lowerUpperBoundsRegret}

@@

\begin{lemma}[Regret decomposition]\label{lem:2:RegretDecomposition}
    The (mean) regret $R_T^{\cA}$ can be written like
    \begin{equation}
        R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ].
    \end{equation}
\end{lemma}

\begin{proof}\label{proof:2:RegretDecomposition}
    Thanks to the tower rule, we had
    $\E \left[ \sum_{t=1}^T r_{A(t)}(t) \right] = \sum_{k=1}^K \mu_k \E \left[ N_k(T) \right]$, thus
    \[ R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ]. \]
    FIXME more details?
\end{proof}

@@

- [Lai and Robbins] lower-bound in $\Omega(\log(T))$
- Worst case lower-bound in $\Omega(\sqrt{T})$
- Adversarial lower-bound in $\Omega(\sqrt{T})$ (also useful for piece-wise stationary models)

@@

- useful decomposition, with number of samples (see Claire's PhD)
- different kinds of bounds, problem-dependent with $\bigO{C \log(T)}$ with $C$ a measure of the problem complexity, or problem-independent with $\bigO{\sqrt{KT}}$ ?
- tools to show regret upper-bounds, concentration inequalities,
- short proof of UCB or klUCB? just to start to show the tools

@@

- Best Arm Identification?
- Quantile regret or just histogram of regrets
- Worst case regret (max $R_T^{r}$ for $r$ index of Monte-Carlo simulation?)
- Fairness

>>>>

\begin{figure}[h!]  % [htbp]
    \centering
    \includegraphics[width=0.75\linewidth]{2-Chapters/2-Chapter/Images/example_of_a_5_arm_bandit_problem__step100.png}
    \caption{Screenshot of the demonstration, after the end of the game, after $t=T=100$ steps.  See \href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}}
    \label{fig:2:example_of_a_5_arm_bandit_problem__step100}
\end{figure}

@@

In the example above, we consider Bernoulli distributions, but other distributions have been studied in the literature.
From now on until the last chapter of this thesis, we only consider stochastic rewards. The piece-wise stationary model is studied in Chapter~\ref{chapter:6}.
%
% Everything is implemented and documented at
% https://smpybandits.github.io/docs/Arms.html
We also focus only on real-valued rewards, meaning that $r_k(t)\in\R$ for all $k$ and $t$.

An important hypothesis is whether rewards are bounded or not,
and whether the player knows if they are bounded or not before starting the bandit game.
Moreover, if rewards are known to be bounded, let say in an interval $[a,b]$, another important hypothesis is whether the player knows the values of $a$ and $b$.
%
Intuitively, the bandit game is easier if the player knows the support of the distributions, and we restrict to this case in all the thesis.
Most of the algorithms proposed in the literature follow this hypothesis as well.
%
The mostly used
infinitely supported distributions are Gaussian, exponential and Poisson,
while in the literature, the Bernoulli distribution is the most common case of finitely supported distributions.
Continuous distributions with finite support also included truncated versions of infinitely supported distributions, in particular truncated Gaussian are used for numerical experiments in lots of research articles.

@@

Given a measure $\lambda$, that is usually the natural Lebesgue measure on $\R^d$, an exponential family of probability distributions is defined as the distributions whose density (relative to $\lambda$) have the following form:
\begin{equation}\label{eq:2:exponentialFamily}
    \Pr(x | \lambda) = h(x) \exp \let( \eta^T T(x) - A(\eta) \right),
\end{equation}
for a parameter vector $\eta$ (the canonical parameter), and given functions $T$ (the sufficient statistic) and $h$.
$A(\eta)$ is the cumulant function, and it is entirely determined by the $\eta$, $h$, $T$,
$A(\eta) = \log \left( \int h(x) \exp(\eta^T T(x)) \lambda(\mathrm{d} x) \right)$.
%
The natural parameter space is the set of values of $\eta$ such that this integral $A(\eta)$ is finite,
and usually the literature focusses on regular and minimal exponential families (when the natural parameter space is an non-empty open set in $\R$).

-- An important result on exponential families is that a distribution in such family is entirely characterized by its parameter $\eta$.
That is why such one-dimensional distributions are entirely characterized by their mean, as $\mu = \E_{\eta}[T(X)]$.

-- A second important result is a simplified form for the Kullback-Leibler divergence \cite{KullbackLeibler51}, for two distributions lying in the same exponential family.
The KL divergence between two distributions $d_1=\Pr(x|\eta_1)$ and $d_2=\Pr(x|\eta_2)$ is defined as
\begin{equation}\label{eq:2:Kullback-LeiblerDivergenceExpFamily1}
    \KL\left(\Pr(x|\eta_1), \Pr(x|\eta_2)\right) = \int \Pr(x|\eta_1) \log(\frac{\Pr(x|\eta_1)}{\Pr(x|\eta_1)}) \lambda(\mathrm{d}x) = \E_{d_1}[ \log(\frac{\Pr(x|\eta_1)}{\Pr(x|\eta_1)}) ].
\end{equation}
%
So this KL divergence can be written $\KL( \eta_1, \eta_2 ) = \KL(\Pr(x|\eta_1), \Pr(x|\eta_2))$, and it can be simplified to an expression using only the two parameters $\eta_1$ and $\eta_2$,
and $\mu_1 = \E_{\eta_1}[T(X)]$ the mean of distribution $d_1$:
%
\begin{equation}\label{eq:2:Kullback-LeiblerDivergenceExpFamily2}
    \KL\left( \eta_1, \eta_2 \right) = \E_{d_1}[ (\eta_1 - \eta_2)^T T(X) ] - A(\eta_1) + A(\eta_2) = (\eta_1 - \eta_2)^T \mu_1 - A(\eta_1) + A(\eta_2),
\end{equation}

@@

\paragraph{Other kind of hypothesis: sub-Gaussian and sub-Bernoulli distributions.}
%
A lot of research works considers rewards distributions that are not Gaussian but sub-Gaussian, of a known variance, for instance $1/4$.
For instance, bounded distributions on $[0,1]$ are known to be $1/4$ sub-Gaussian, and this fact is used for instance in \cite{Maillard2018GLR}.
It means that their moment generating function is dominated by that of a Gaussian distribution with the same mean and a variance $1/4$.
In Chapter~\ref{chapter:6}, we instead consider sub-Bernoulli distributions, formally introduced in Definition~\ref{def:6:subBernoulliDistributions}.
%
Such hypothesis is also proposed for other families of distributions, for instance \cite{KimTewari2019} analyses their Follow-the-Perturbed-Leader algorithm for perturbations following any sub-Weibull distribution (sub-Weibull distributions generalize both sub-Gaussian and sub-Exponential distributions).

@@

\paragraph{Applications for Opportunistic Spectrum Access.}
% Detail the previous work from our team SCEE on bandits + OSA : Wassim, Navik
%
Previous works of our SCEE team showed that MAB can be used to model the problem of spectrum access for a secondary user accessing a licensed spectrum.
In this model, arms represent a finite set of orthogonal channels, \ie, different frequency bands in a licensed spectrum.
In the model with sensing, the rewards $r_k(t)$ represents the feedback obtained by the CR-equipped device after sensing the channel $k$ at time $t$.
A reward of $r_k(t) = 1$ indicates that no Primary User was sensed, while a reward of $r_k(t)=0$ indicates that the channel $k$ is busy at time $t$ and no uplink message can be sent.
%
This model was first introduced by Wassim Jouini during his PhD thesis,
in \cite{Jouini09} and later studied in both \cite{Jouini10,Jouini12}.
Proof-of-concepts using real-world radio hardware was first propsed in \cite{MoyWSR2014,RobertSDR2014}.
In a second PhD thesis supervised by Christophe Moy, Navikkumar Modi studied the impact
of using MAB algorithms to optimize channel selection
on the battery life of a wireless device.
On the one hand, running a MAB algorithm such as UCB-like algorithms was proved to be useful and can bring significant improvement in terms of successful transmission rates, directly increasing the battery life of the device.
On the other hand, classical MAB algorithms tend to switch arms a lot of times, especially in the beginning of the learning process, and this induces a lot of dynamical hardware reconfiguration for the wireless device, as selecting a different channel requires a change in the radio hardware used by the device.
Each hardware reconfiguration costs energy for the device, and quickly switching algorithms will lead to a reduction of the battery life.
The tradeoff between the two aspects is studied empirically in
\cite{modiDemo2016}.
%
Another interesting works is \cite{Modi17QoS}.

\TODOL{FIXME Finir ce blabla rapidement ? Ou alors ce sera déjà fait en intro/chapitre1 ??}

% ----------------------------------------------------------------------------
\section{Regret: definition, decomposition, and lower bounds}
\label{sec:2:lowerUpperBoundsRegret}

@@

\begin{lemma}[Regret decomposition]\label{lem:2:RegretDecomposition}
    The (mean) regret $R_T^{\cA}$ can be written like
    \begin{equation}
        R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ].
    \end{equation}
\end{lemma}
%
\begin{proof}\label{proof:2:RegretDecomposition}
    Thanks to the tower rule, we had
    $\E \left[ \sum_{t=1}^T r_{A(t)}(t) \right] = \sum_{k=1}^K \mu_k \E \left[ N_k(T) \right]$, thus
    \[ R_T^{\cA} = \sum_{k=1}^K \Delta_k \E[ N_k(T) ]. \]
    % FIXME more details?
\end{proof}

@@

\begin{itemize}
    \item
    - [Lai and Robbins] lower-bound in $\Omega(\log(T))$
    \item
    - Worst case lower-bound in $\Omega(\sqrt{T})$
    \item
    - Adversarial lower-bound in $\Omega(\sqrt{T})$ (also useful for piece-wise stationary models)
\end{itemize}

@@

\begin{itemize}
    \item
    - different kinds of bounds, problem-dependent with $\bigO{C \log(T)}$ with $C$ a measure of the problem complexity, or problem-independent with $\bigO{\sqrt{KT}}$ ?
    \item
    - tools to show regret upper-bounds, concentration inequalities?
    FIXME non j'ai pas la place
    \item
    - short proof of UCB or klUCB? just to start to show the tools
    FIXME non j'ai pas la place
\end{itemize}

@@

- Best Arm Identification?
- Quantile regret or just histogram of regrets
- Worst case regret (max $R_T^{r}$ for $r$ index of Monte-Carlo simulation?)
- Fairness ?
- Switching cost? \cite{modiDemo2016} \cite{Koren17}
